General constraints on magic distillation protocols from statistical mechanics

### 2021-06-06 magic_paper.tex
Minor typos. Done for day.

The $\alpha \rightarrow 1$ limit of the Renyi entropy is not well-defined, however the Renyi entropy description shows that this diverges to $-\infty$ due to the Lorenz curve $L_\rho(x)$ no longer being bounded by $1$. Alternatively, we can write the divergence as $-\epsilon \log \sum_{\x} |W_\rho (\x)|^{1+\epsilon}$ with $\epsilon \rightarrow 0^+$, which leads to a second new description of mana: \emph{it is the rate of divergence of the Renyi-entropy} as we approach the asymptotic Shannon entropy $H(\w)$. We also note the recent observation \ddd{[CITE Renyi Entropy and Free Energy by John C. Baez]}that in standard statistical mechanics that $H_\alpha(\p)$ can be related to the ``$q$--derivative'' of free energy. Therefore, we might speculate if the presence of magic in Clifford circuits admits a rigorous interpretation in terms of a phase transition signalled by a first-order divergence in the Wigner free energy. However since variation in $\alpha$ is no longer smooth ($\alpha$ is restricted to a dense set of rational values) technical obstacles exist and we do not pursue this speculative line here, but just note that other works exist in which computational transitions are considered in relation to phase transitions [Need to cite].

>>>>

The $\alpha \rightarrow 1$ limit of the Renyi entropy is not well-defined, however the Renyi entropy description shows that this diverges to $-\infty$ due to the Lorenz curve $L_\rho(x)$ no longer being bounded by $1$. Alternatively, we can write the divergence of the entropy as $-\epsilon^{-1} \log \sum_{\x} |W_\rho (\x)|^{1+\epsilon}$ with $\epsilon \rightarrow 0^+$, which leads to a second new description of mana: \emph{it is the rate of divergence of the Renyi-entropy} as we approach the limiting Shannon entropy $H(\w)$. We also note the recent observation \ddd{[CITE Renyi Entropy and Free Energy by John C. Baez]}that in standard statistical mechanics that $H_\alpha(\p)$ can be related to the ``$q$--derivative'' of free energy. Therefore, we might speculate if the presence of magic in Clifford circuits admits a rigorous interpretation in terms of a phase transition signalled by a first-order divergence in the Wigner free energy. However since variation in $\alpha$ is no longer smooth ($\alpha$ is restricted to a dense set of rational values) technical obstacles exist and we do not pursue this speculative line here, but just note that other works exist in which computational transitions are considered in relation to phase transitions \ddd{[Need to cite]}.
